---
title: "Turkey University Analysis"
format:
  html:
    code-fold: true
    code-summary: "Show the code"
---

Welcome to my project page. In this page, you will see various information about my project. These include basic explanation about the YOK Atlas data (rows, columns, etc.), reason of choice of this data and finally some analysis with this data. Main aim in this project is to represent the status of universities in Turkey and examine many aspects of them.

# 1. Project Overview and Scope

This project is based on the universities in Turkey. Turkey has many universities but every one of them carry different attributes like scores, city, region, etc. We can clearly say that it is not possible to grasp all this information with the raw data only so this project will represent many aspect of this data. These aspects include basic statistical information, type of the university (Private or State), region, detailed information of the universities (the number of exchange students, assistant professor etc.) and so on. By this way, this all information can be seen visually, filtered and grasp easily.

# 2. Data

Data consists of three main titles: basic information of universities, score information (i.e. the number of exchange students), and department information.

***a. Basic Information***

Here you will see the first 10 rows of basic information of the universities:

```{r}
suppressPackageStartupMessages(library(thestats))
library(thestats)
library(knitr)

kable(head(list_uni(), 10))
```

This data gives 6 information: id, name, type, city, region, and website of universities.

***b. Numerical Information***

This data shows many numerical attributes of the universities (205 columns). For the explanation about the attributes, reader can visit [here](https://github.com/analyticsresearchlab/thestats/blob/main/docs/Table3.md).

```{r}
suppressPackageStartupMessages(library(thestats))
library(thestats)
library(knitr)

kable(head(list_score(department_names = "all"), 10))
```

***c. Department Information***

This dataset contains information based on the department in the universities. These information contain id, year, type, program code, university name and faculty.

```{r}
suppressPackageStartupMessages(library(thestats))
library(thestats)
library(knitr)
kable(head(list_dept(department_names = "all"), 10))
```

## 2.1 Data Source

This data is prepared by YOK and for analysis purposes, it is scraped and organized in an R package with the name [thestats](https://github.com/analyticsresearchlab/thestats). You can visit the link to learn more about it.

## 2.2 General Information About Data

As shown above, data classified into three main titles: basic, score and department information. From these data, many summaries and knowledge can be extracted like the number of private and state universities based on the region, the number of universities that has the specific department like statistics from the department information data and so on.

These data contain many columns and these columns can be examined in the given links above.

## 2.3 Reason of Choice

Universities are so important for a country in many ways but this comes with a really serious problem: the number of universities in Turkey. This project's main aim is to analyse these all universities in many ways like total number of them, geographical (regional) based knowledge etc. After this analysis, reader can learn the student's various behavior about the universities. One of the most important thing is that people's ideas can change based on the geographical status of the university like department choice, university type, choice between the priority of university or department etc. For example, one student could want to study engineering but only can go to some universities in East. We can learn many detailed information quickly like in the example.

# 3. Analysis

## 3.1 Exploratory Data Analysis

```{r}
suppressPackageStartupMessages(library(thestats))
library(thestats)
library(dplyr)
library(ggplot2)

df <- list_uni() %>% count(type)
private <- df[1, "n"]
state <- df[2, "n"]

cat("The number of universities in Turkey is", length(unique(list_uni()$university)), ".", private, "of them are Private and", state, "of them are State university.")

cat("After learning the number of private and state universities, the distribution of the type of universities grouped by regions can be interesting to analyze.")

private_uni <- list_uni()[list_uni()$type == "Private", ] %>%
  group_by(region) %>%
  summarize(num_universities = n_distinct(university))

state_uni <- list_uni()[list_uni()$type == "State", ] %>%
  group_by(region) %>%
  summarize(num_universities = n_distinct(university))

region_counts <- list_uni() %>%
  group_by(region) %>%
  summarize(num_universities = n_distinct(university))

ggplot(private_uni, aes(x = region, y = num_universities)) +
  geom_bar(stat = "identity", fill = "#ed4c07") +
  labs(x = "Region", y = "Number of Private Universities") +
  ggtitle("Number of Private Universities by Region") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(state_uni, aes(x = region, y = num_universities)) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(x = "Region", y = "Number of State Universities") +
  ggtitle("Number of State Universities by Region") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

To see the number of universities based on the region, below chart will be helpful. As can be seen, the maximum number of universities is in Marmara region and continues with Central Anatolia, Black Sea etc.

```{r}
suppressPackageStartupMessages(library(thestats))
library(thestats)
library(ggplot2)

region_counts <- list_uni() %>%
  group_by(region) %>%
  summarize(num_universities = n_distinct(university))

ggplot(region_counts, aes(x = region, y = num_universities)) +
  geom_bar(stat = "identity", fill = "#ed4c07") +
  labs(x = "Region", y = "Number of Universities") +
  ggtitle("Number of Universities by Region") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

With the help of the above chart can say that for Marmara region and Central Anatolia, special attention should be taken because in these regions the number of students will be much more when compared with others. These actions include dormitories, libraries, social activities and such.

## 3.2 A Special Attention to Marmara Region

A special attention is taken to Marmara and the top 20 faculties are shown above in the region. 

```{r}

library(knitr)

faculty_group <- list_score(region_names = "Marmara", department_names = "all") %>% group_by(faculty) %>% 
  summarize(num_faculties = length(faculty))

faculty_numbers <- faculty_group[order(faculty_group$num_faculties, decreasing = TRUE),]

kable(head(faculty_numbers, 20))
```

## 3.3 Students' Choice Order is Important

Every student choices universities in a certain order and based on the point obtained in the university exam, they are accepted to the university. Below, I will explore the top 1st, 2nd and 3rd choice universities. This information can be interesting and helpful to other universities that seeks a way to develop and popularitiy or some other interesting results can be obtained from this analysis.

```{r}

library(knitr)

sorted_1st_choice <- head(list_score(department_names = "all")[order(list_score(department_names = "all")$X141, decreasing = TRUE),], 10)
row.names(sorted_1st_choice) <- NULL

sorted_2nd_choice <- head(list_score(department_names = "all")[order(list_score(department_names = "all")$X142, decreasing = TRUE),], 10)
row.names(sorted_2nd_choice) <- NULL

sorted_3rd_choice <- head(list_score(department_names = "all")[order(list_score(department_names = "all")$X143, decreasing = TRUE),], 10)
row.names(sorted_3rd_choice) <- NULL

filter_exp_1 <- c("university", "faculty", "department", "X141")
filter_exp_2 <- c("university", "faculty", "department", "X142")
filter_exp_3 <- c("university", "faculty", "department", "X143")

first_filtered <- sorted_1st_choice[filter_exp_1]
colnames(first_filtered) <- c("university", "faculty", "department", "number_of_1st_choice")

cat("1st Choice Universities: ")

kable(first_filtered)

second_filtered <- sorted_2nd_choice[filter_exp_2]
colnames(second_filtered) <- c("university", "faculty", "department", "number_of_2nd_choice")

cat("2nd Choice Universities: ")

kable(second_filtered)


third_filtered <- sorted_3rd_choice[filter_exp_3]
colnames(third_filtered) <- c("university", "faculty", "department", "number_of_3rd_choice")

cat("3rd Choice Universities: ")

kable(third_filtered)
```

As can be seen above table, Anadolu University Distance Education is the first university for the 1st, 2nd and 3rd choice. For a more reliable analysis, Anadolu University Distance Education faculty is removed from the list and other universities examined.

```{r}

library(knitr)

orig_data <- list_score(department_names = "all")
anad_univ_removed <- orig_data[orig_data$university!="Anadolu University",]

sorted_1st_choice <- head(anad_univ_removed[order(anad_univ_removed$X141, decreasing = TRUE),], 10)
row.names(sorted_1st_choice) <- NULL

sorted_2nd_choice <- head(anad_univ_removed[order(anad_univ_removed$X142, decreasing = TRUE),], 10)
row.names(sorted_2nd_choice) <- NULL

sorted_3rd_choice <- head(anad_univ_removed[order(anad_univ_removed$X143, decreasing = TRUE),], 10)
row.names(sorted_3rd_choice) <- NULL

filter_exp_1 <- c("university", "faculty", "department", "X141")
filter_exp_2 <- c("university", "faculty", "department", "X142")
filter_exp_3 <- c("university", "faculty", "department", "X143")

first_filtered <- sorted_1st_choice[filter_exp_1]
colnames(first_filtered) <- c("university", "faculty", "department", "number_of_1st_choice")

cat("1st Choice Without Anadolu University: ")

kable(first_filtered)

second_filtered <- sorted_2nd_choice[filter_exp_2]
colnames(second_filtered) <- c("university", "faculty", "department", "number_of_2nd_choice")

cat("2nd Choice Without Anadolu University: ")

kable(second_filtered)

third_filtered <- sorted_3rd_choice[filter_exp_3]
colnames(third_filtered) <- c("university", "faculty", "department", "number_of_3rd_choice")

cat("3rd Choice Without Anadolu University: ")

kable(third_filtered)
```

## 3.4 Highest Score of Placed Students

Highest score of placed Students' plays an important role in the success and the popularitiy of the university. Below I will show an analysis about the most successful students' based on the university. By examining this, we will get a chance to learn the most successful universities in Turkey.

```{r}

library(knitr)

orig_data <- list_score(department_names = "all")
orig_data <- orig_data[orig_data$X16 != "---",]

orig_data$X16 <- as.numeric(gsub(",", ".", orig_data$X16))

highest_score_student <- orig_data %>%
  group_by(university) %>%
  summarize(average_score = mean(X16))

kable(head(highest_score_student[order(highest_score_student$average_score, decreasing = TRUE),], 20))

```

## 3.5 Number of Professors in Universities

While examining the status of a university, the number of professors and associate professors is so important. Professors are so experienced and students and other lecturers can learn so much information in a short time. By examining the numbers, we can contruct a relationship with the success and the popularity of the universities.

```{r}

library(knitr)

orig_data <- list_score(department_names = "all")
orig_data$prof_plus_assoc_prof <- orig_data$X188 + orig_data$X189

highest_prof_num <- orig_data %>%
  group_by(university) %>%
  summarize(average_prof = mean(prof_plus_assoc_prof))

kable(head(highest_prof_num[order(highest_prof_num$average_prof, decreasing = TRUE),], 10))

cat("The Worst Universities Based on the Number of Professors and Associate Professors: ")

kable(head(highest_prof_num[order(highest_prof_num$average_prof, decreasing = FALSE),], 10))

```

# 4 Results and Last Discussions

Universities are so important in many aspects. Any doctor, engineer, teacher, lecturer etc. comes from these organisations. Analyzing the status of these is so helpful to understand many things at a first glance. This work is just a beginning of such a complete work. Data contains almost 200 different information and here the most important information is examined. Many organizations and regulations and much more can be done with these informations. I hope this small work helped to understand the different aspects of our universities.